Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Remove unnecessary statement that A is constant
Reorganize paragraph discussin solutions for the modal collapse

\begin{small}
\begin{prooftree}
\AXC{Theorem 2}\dashedLine
\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}
\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
\end{prooftree}
\end{small}

@@

In G\"odel's ontological proof, we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being. The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas. Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being god-like or at least to a restricted collection of properties. Several solutions to the problem of modal collapse have been proposed.

Anderson's solution \citep{and90} modifies the definitions of God-like being and essence, and eliminates half of an axiom. This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}. Another solution involving more substantial modifications is that of Bj{\o}rdal \citep{bjordal,fuhrmann}.

On another track, Fitting has argued that greater care has to be taken with the semantics of higher-order modal logics. Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional. Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].

Anderson \citep{and90}[p. 292] and Sobel \citep{sobel2}[p. 133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.

A reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.

@@

\begin{small}
\begin{prooftree}
\AXC{Theorem 2}\dashedLine
\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$}
\UIC{$\all y.[G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
\UIC{$\all y.[G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
\UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine
\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
\end{prooftree}
\end{small}

>>>>

\begin{small}
\begin{prooftree}
\AXC{Theorem 2}\dashedLine
\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}
\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
%\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
\end{prooftree}
\end{small}

@@

In G\"odel's ontological proof, we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being. The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas. Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being god-like or at least to a restricted collection of properties. Several solutions to the problem of modal collapse have been proposed.
Fitting, for instance, has argued that greater care has to be taken with the semantics of higher-order modal logics. Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional. Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10]. Koon \citep{koon}, Anderson \citep{and90}[p. 292] and Sobel \citep{sobel2}[p. 133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate. Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next section.

%A reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.

@@

\begin{small}
\begin{prooftree}
\AXC{Theorem 2}\dashedLine
\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$}
\UIC{$\all y.[G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
%\UIC{$\all y.[G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
\UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine
\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
\end{prooftree}
\end{small}
